Phyloseq Preprocessing

### 2022-03-12 Rmarkdown/analysis/Response_to_Reviewers.Rmd

---
title: "Response to Reviewers"
author: "Kristen James"
date: "3/7/2022"
output: html_document
---

This markdown was created in response to the Reviewer's comments during the peer-review process of our manuscript submitted to Nutrients. The workflow is borrowed in large part from "Characteristics_Diet_CardiometabolicRisk_211111.Rmd" with subtle changes.

Load libraries

```{r Libraries, warning=FALSE, message=FALSE}
library(plyr)
library(tableone)
```

Load data

```{r Load data}
phen <- readRDS("../../data/processed/phenotype/Phenotypes_211108.rds")

# Complete TMAO data only
phen <- phen[complete.cases(phen$tmao == TRUE),]

# Get rid of NA ethnicity category
phen$ethnicity <- droplevels(phen$ethnicity)

# Factor and recode sex
phen$sex <- as.factor(phen$sex)
phen$sex <- plyr::revalue(phen$sex, c("1" = "Male", "2" = "Female"))

# Make TMAO log variable
phen$tmao_log <- log(phen$tmao)

```

# Reviewer 1

Provide the number of participants in each age-sex-BMI bin to answer Reviewer 1's comment about statistical power.

```{r}
table(phen$bin_number)
```

Reviewer 1 didn't understand that the relationship between TMAO and age differed by sex. We will make it more clear by stating the restults of the statistical test.

```{r sex*age interaction}
summary(lm(tmao_log ~ sex*age  + cystatinc_bd1, phen))
```

# Reviewer 2

## Fiber Comment
Reviewer 2 asked why we did not include fiber in the analysis. We will add total fiber intake from foods and supplements (fibe_tnfs) to our analysis.

First, lets set up the dataframe. This code is borrowed from "Characteristics_Diet_CardiometabolicRisk_211111.Rmd".

```{r Load data and preprocess}
phen <- readRDS("../../data/processed/phenotype/Phenotypes_211108.rds")

# Complete TMAO data only
phen <- phen[complete.cases(phen$tmao == TRUE),]

# Get rid of NA ethnicity category
phen$ethnicity <- droplevels(phen$ethnicity)

# Factor and recode sex
phen$sex <- as.factor(phen$sex)
phen$sex <- plyr::revalue(phen$sex, c("1" = "Male", "2" = "Female"))

# Make TMAO log variable
phen$tmao_log <- log(phen$tmao)

# Convert crp units
phen$crp_bd1.mgdL <- phen$crp_bd1 * 0.0001
```

Create quantile and tertile classifiers. Transform data to comply to normality (Shapiro-Wilk test W statistic > 0.95).

```{r Quantile and Tertile TMAO}
# Quantile
tmao_quantile <- quantile(phen$tmao)
phen$tmao_quantile <- ifelse(phen$tmao <= tmao_quantile[2], "Quantile1",
                             ifelse(phen$tmao > tmao_quantile[2] & phen$tmao <= tmao_quantile[3], "Quantile2",
                                   ifelse(phen$tmao > tmao_quantile[3] & phen$tmao <= tmao_quantile[4], "Quantile3", "Quantile4")))

# Check it
table(phen$tmao_quantile)
phen$tmao_quantile <- as.factor(phen$tmao_quantile)
str(phen$tmao_quantile)

# Tertile
tmao_tertile <- quantile(phen$tmao, c(0:3/3))
phen$tmao_tertile <- ifelse(phen$tmao <= tmao_tertile[2], "Tertile1",
                             ifelse(phen$tmao > tmao_tertile[2] & phen$tmao <= tmao_tertile[3], "Tertile2", "Tertile3"))
# Factor it
phen$tmao_tertile <- factor(phen$tmao_tertile)
table(phen$tmao_tertile)

# Print tertiles for manuscript
tmao_tertile

# Checked all variables for normality via shapiro.test() in console. TG, and glc ere the only ones that didn't pass.
phen$glc_bd1_ln <- log(phen$glc_bd1)
phen$tg_bd1_ln <- log(phen$tg_bd1)
phen$crp_bd1_ln <- log(phen$crp_bd1)
phen$tnfa_bd1_ln <- log(phen$tnfa_bd1)
phen$il6_bd1_ln <- log(phen$il6_bd1)
phen$insulin_bd1_ln <- log(phen$insulin_bd1)
```

Load the diet data.

```{r Load ASA24 averaged data}
food <- readRDS(file = "../../data/processed/asa24/ASA24_meanIntake_withTransformations_211111.rds")

# Combine food with phenotype data
full <- merge(phen, food, by = 0)
row.names(full) <- full[,"Row.names"]
full <- subset(full, select = -c(Row.names))

# Make ASA24 choline log variable
full$choln_tnfs_log <- log(full$choln_tnfs)

# Make ASA24 fiber log variable
shapiro.test(log(full$fibe_tnfs))
full$fibe_tnfs_log <- log(full$fibe_tnfs)
```

Make a list of foods and covariate variables to keep. Add fiber this time!

```{r Foods to keep, results='hide'}
transformed_keep_tbl <- c("sex", 
#"sex_numeric",
"kcal_tnfs", 
"age", 
"cystatinc_bd1", 
"ethnicity",
"tmao", 
"tmao_log", 
#"TMAO.0",
#"TMAO.30",
#"TMAO.180",
#"TMAO.360",
#"lnTMAO.0",
#"lnTMAO.30",
#"lnTMAO.180",
#"lnTMAO.360",
#"TMAO_0_30",
#"TMAO_0_180",
#"TMAO_0_360",
#"Ind.SO4",
#"p.Cresol.SO4",
#"TMAO_quart_num",
#"T.pf_mps_totals_sqrRt",
"T.pf_meat_totals_sqrRt",
"T.pf_curedmeat_totals_sqrRt",
"T.pf_poult_totals_sqrRt",
"T.pf_seafd_hi_totals_Binary",
"T.pf_seafd_low_totals_Binary",
"T.pf_eggs_totals_sqrRt",
"T.pf_nutsds_totals_sqrRt",
"T.pf_legumes_totals_sqrRt",
"T.pf_total_totals_natLog",
#"T.pf_soy_totals_Binary",
"T.v_drkgr_totals_sqrRt",
"T.v_redor_total_totals_sqrRt",
#"T.v_redor_tomato_totals_sqrRt",
#"T.v_redor_other_totals_sqrRt",
"T.v_starchy_total_totals_sqrRt",
#"T.v_starchy_potato_totals_sqrRt",
#"T.v_starchy_other_totals_sqrRt",
"T.v_other_totals_sqrRt",
"T.v_total_totals_sqrRt",
#"T.v_legumes_total_sqrRt",
"T.g_whole_totals_sqrRt",
"T.g_refined_totals_sqrRt",
"g_total_totals",
"T.d_total_totals_sqrRt",
"T.d_milk_totals_sqrRt",
"T.d_cheese_totals_sqrRt",
"T.d_yogurt_totals_Binary",
"choln_tnfs_log",
"hei_asa24_totalscore",
"fibe_tnfs_log"
)
```

Now run the multiple linear regression. Note, we will not include ethnicity in these regressions because although TMAO levels significantly differ between ethnic groups, there was no relationship between TMAO and ethnicity. Additionally, diet patterns differ by ethnicity so we could be over adjusting.

```{r TMAO ~ food - LCMS BD1 - lm 356 ppl, results='hide'}
# Make a smaller df from master (the old data)
full_sub <- full[, colnames(full) %in% transformed_keep_tbl]
food <- subset(full_sub, select = -c(sex, kcal_tnfs, age, cystatinc_bd1, ethnicity, tmao, tmao_log))
n <- ncol(food)

# LCMS method
# Run linear model
my_lm <- lapply(1:n, function(x) lm(tmao_log ~ food[,x] + sex*age + cystatinc_bd1 + kcal_tnfs, full_sub )) # include kcal

# Organize results
my_lm2 <- sapply(my_lm, function(x){summary(x)$adj.r.squared})
my_lm2.5 <- sapply(my_lm, function(x){summary(x)$coefficients[2,1]}) # [food row, estimate] corresponds to food B 
my_lm2.6 <- sapply(my_lm, function(x){summary(x)$coefficients[2,2]}) # [food row, Std.Error] corresponds to food Std. Error 
my_lm2.7 <- sapply(my_lm, function(x){summary(x)$coefficients[2,3]}) # [food row, T value] corresponds to food t value
my_lm3 <- sapply(my_lm, function(x){summary(x)$coefficients[2,4]}) # [food row, P] corresponds to food P value
my_lm356_lcms <- cbind(my_lm2, my_lm2.5, my_lm2.6, my_lm2.7, my_lm3)
colnames(my_lm356_lcms) <- c("adj.r.squared", "estimate", "Std.Error", "T.value","P")
row.names(my_lm356_lcms) <- colnames(food)
res_food <- as.data.frame(my_lm356_lcms)

# BH correction
res_food$p.adj <- stats::p.adjust(res_food$P, method = "BH", n = n) 
```

Save with the extension "_withFiber"

```{r save food linear regression table}
# Save table
write.csv(res_food, "../../Outputs/tables/Diet/MultipleLinReg_TMAO_LCMS_n356_ASA24_withFiber.csv")
```

We also need to add fiber to our food by TMAO tertile table.

```{r Food Intake by TMAO tertile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("pf_meat_totals", 
          "pf_curedmeat_totals",
          "pf_poult_totals",
          "pf_seafd_hi_totals",
          "pf_seafd_low_totals",
          "pf_eggs_totals",
          "pf_nutsds_totals",
          "pf_legumes_totals",
          "pf_total_totals",
          #"pf_soy_totals_Binary",
          "v_drkgr_totals",
          "v_redor_total_totals",
          #"v_redor_tomato_totals",
          #"v_redor_other_totals",
          "v_starchy_total_totals",
          #"v_starchy_potato_totals",
          #"v_starchy_other_totals",
          "v_other_totals",
          "v_total_totals",
          #"v_legumes_total",
          "g_whole_totals",
          "g_refined_totals",
          "g_total_totals",
          "d_total_totals",
          "d_milk_totals",
          "d_cheese_totals",
          "d_yogurt_totals",
          "hei_asa24_totalscore",
          "choln_tnfs",
          "fibe_tnfs", # fiber added!
          "age",
          "ethnicity" ,
          "bmi_final",
          "tmao",
          "choline",
          "betaine",
          "carnitine")
table_food.ft <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full[full$sex == "Female",],
                                    factorVars = factorVars)
table_food.ft

# TMAO table by age bin - males
table_food.mt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full[full$sex == "Male",],
                                    factorVars = factorVars)
table_food.mt

# TMAO table by sex
table_food.allt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full,
                                    factorVars = factorVars)
table_food.allt

# Do BH correction
food_allt_ContTable <- attr(table_food.allt$ContTable, "pValues")
n <- length(food_allt_ContTable$pNormal)
# BH correction
food_allt_ContTable$padj <- stats::p.adjust(food_allt_ContTable$pNormal, method = "BH", n = n) 
food_allt_ContTable
```

Save with the "_withFiber" extension.

```{r Save food tertile tables}
table_save.ft <- print(table_food.ft, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.ft, file = "../../Outputs/tables/ParticipantChar/Food_Female_byTMAOTertile_withFiber.csv")

table_save.mt <- print(table_food.mt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mt, file = "../../Outputs/tables/ParticipantChar/Food_Male_byTMAOTertile_withFiber.csv")

table_save.allt <- print(table_food.allt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allt, file = "../../Outputs/tables/ParticipantChar/Food_All_byTMAOTertile_withFiber.csv")

write.csv(food_allt_ContTable, file = "../../Outputs/tables/ParticipantChar/Food_All_byTMAOTertile_PadjTable_withFiber.csv")
```

Reviewer 2 asked us to clarify what transformations were used in the endothelial analysis. Ultimately, we used AI in Table 1 and  AI75 in the linear regression analysis. This was because when we use AI75 in the descriptive table, the mean values are negative and positive, which is confusing/not intuitive. The following chunk demonstrates this point.

```{r Participant characterisitics by TMAO tertile}
# TMAO table by tmao_tertile - females
factorVars <- c("age_cat", "tmao_tertile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "insulin_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.ft <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen[phen$sex == "Female",],
                                    factorVars = factorVars)
table_tmao.ft

# TMAO table by tmao_tertile - males
factorVars <- c("age_cat", "tmao_tertile")
vars <- c("age", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "insulin_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.mt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen[phen$sex == "Male",], # note 6103 and 7098 have high il6
                                    factorVars = factorVars)
table_tmao.mt

# TMAO table by tmao_tertile - all, add sex
factorVars <- c("tmao_tertile", "sex")
vars <- c("age", "sex", "ethnicity" , "bmi_final", "waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1", "glc_bd1", "insulin_bd1", "tnfa_bd1", "il6_bd1", "crp_bd1.mgdL", "cystatinc_bd1", "tmao", "choline", "betaine", "carnitine")
table_tmao.allt <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = phen,
                                    factorVars = factorVars)
table_tmao.allt

# Do BH correction for continuous variables
allt_ContTable <- attr(table_tmao.allt$ContTable, "pValues")
n <- length(allt_ContTable$pNormal)
# BH correction
allt_ContTable$padj <- stats::p.adjust(allt_ContTable$pNormal, method = "BH", n = n) 
allt_ContTable

# Do BH correction for categorical variables
allt_CatTable <- attr(table_tmao.allt$CatTable, "pValues")
n <- 23
# BH correction
allt_CatTable$padj <- stats::p.adjust(allt_CatTable$pApprox, method = "BH", n = n) 
allt_CatTable
```

Save with "_al75" extension; however, we did not use these tables (we used the AI version originally created).

```{r Save characteristic tertile tables}
table_save.ft <- print(table_tmao.ft, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.ft, file = "../../Outputs/tables/ParticipantChar/Characteristics_Female_byTMAOTertile_al75.csv")

table_save.mt <- print(table_tmao.mt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.mt, file = "../../Outputs/tables/ParticipantChar/Characteristics_Male_byTMAOTertile_al75.csv")

table_save.allt <- print(table_tmao.allt, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
write.csv(table_save.allt, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_byTMAOTertile_al75.csv")

# BH corrections for the all tertile table
write.csv(allt_ContTable, file = "../../Outputs/tables/ParticipantChar/Characteristics_All_byTMAOTertile_Padj_al75.csv")
```

Do some individuals have undiagnosed hyperglycemia or hyperlipidemia? Yes

```{r undiagnosed metabolic disease}
summary(phen$glc_bd1)
summary(phen$tg_bd1)
summary(phen$chol_bd1)

hist(phen$glc_bd1, breaks = 40)
hist(phen$tg_bd1, breaks = 40)
hist(phen$chol_bd1, breaks = 40)
```

### 2022-03-24 Rmarkdown/analysis/Response_to_Reviewers.Rmd

---
title: "Response to Reviewers"
author: "Kristen James"
date: "3/7/2022"
output: html_document
---

This markdown was created in response to the Reviewer's comments during the peer-review process of our manuscript submitted to Nutrients. The workflow is borrowed in large part from "Characteristics_Diet_CardiometabolicRisk_211111.Rmd" with subtle changes.

Load libraries

```{r Libraries, warning=FALSE, message=FALSE}
library(plyr)
library(tableone)
library(ggplot2)
```

@@

```{r}
table(phen$bin_number)
```

Reviewer 1 didn't understand that the relationship between TMAO and age differed by sex. We will make it more clear by stating the restults of the statistical test.

```{r sex*age interaction}
summary(lm(tmao_log ~ sex*age  + cystatinc_bd1, phen))
```

## Reviewer 1 - Response 2
Demonstrate the sex* age interaction in a visual plot for the supplement.
```{r}
colnames(phen)
sd <- sd(phen$tmao_log)

# too messy, point of interaction not obvious
ggplot(phen, aes(x = age, y = tmao_log, group = sex, color = sex)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=tmao_log - sd, ymax=tmao_log+sd),
                width = .2,
                position = position_dodge(0.05)) +
  theme_minimal() +
  ylab("Log of TMAO (uM)") +
  xlab("Age (years)") +
  scale_color_brewer(palette = "Dark2") +
  labs(color = "Sex")

# try a simplified version - this is better
p <- ggplot(phen, aes(x = age, y = tmao_log, group = sex, color = sex)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  ylab("Log of TMAO (uM)") +
  xlab("Age (years)") +
  scale_color_brewer(palette = "Dark2") +
  labs(color = "Sex")

tiff(paste0("../../Outputs/plots/Sex_x_Age_Interaction.tiff"), width = 5, height = 4, units = 'in', res = 300)
p
dev.off()
```

### 2022-03-24 Rmarkdown/analysis/Response_to_Reviewers.Rmd

# Reviewer 1
## Reviewer 1 - Response 1

Provide the number of participants in each age-sex-BMI bin to answer Reviewer 1's comment about statistical power.

```{r}
table(phen$bin_number)
```

Reviewer 1 didn't understand that the relationship between TMAO and age differed by sex. We will make it more clear by stating the results of the statistical test.

@@

```{r sex*age interaction}
summary(lm(tmao_log ~ sex*age  + cystatinc_bd1, phen))
```

## Reviewer 1 - Response 2
Demonstrate the sex* age interaction in a visual plot for the supplement.
```{r}
colnames(phen)
sd <- sd(phen$tmao_log)

# Plot
p <- ggplot(phen, aes(x = age, y = tmao_log, group = sex, color = sex)) +
  geom_point() +
  geom_smooth(method = lm) +
  theme_minimal() +
  ylab("Log of TMAO (uM)") +
  xlab("Age (years)") +
  scale_color_brewer(palette = "Dark2") +
  labs(color = "Sex")

tiff(paste0("../../Outputs/plots/Sex_x_Age_Interaction.tiff"), width = 5, height = 4, units = 'in', res = 300)
p
dev.off()
```

### 2022-03-24 Rmarkdown/preprocessing/Phyloseq_Preprocessing_211104.Rmd
Limit to TMAO analysis

1) Processed QIIME2 inputs in R .rds and .csv formats
2) Phyloseq object with log transformed Biocrates BD1 data [PSOlogBioc]
3) Phyloseq object with TMAO releated phenotypes [PSOtmao]

**Updates:**
On 01-19-2022 I updated the chunks, "Original Data from Mary" and "Make Phyloseq object- specific TMAO variables". When MK was reading my manuscript, she noticed that I have some entries in the taxonomy table that say NA and some that say "g__" or "f__". While these may have different meanings, she suggests simply changing all cases to NA. (Another solution, you could manually differentaite "g__" to "g__ASV1" so that the glomming isn't wrong). I added code (from MK) to consistently change the NAs. Then, I created a new phyloseq object, PSO_tmao_NA. These will be added to the project branch named "update-taxonomy-table" in this GitHub project.

@@

## Biocrates Phyloseq Object

Make a phyloseq object for the log transformed biocrates BD1 data. First, we have to do additional preprocessing.

```{r Biocrates BD1 log transformed phyloseq preprocess}
# Load Mary's original memtadata file... that we converted into a csv in the above chunk. 
md <- readRDS("../../data/processed/microbiome/phyloseq_inputs/FL100_mapping_file_complete_highest-seq-reps_nocontrols_min_withAlphaStatistics_211104.rds")
rownames(md) <- md[,"Description"]

# Load phenotype data HERE 
df <- readRDS("../../data/processed/biocrates/FL100_biocrates_lnTransformed_BD1_211103.rds")
colnames(df[,1:10])

# Trim df to subjects WITH TMAO (primary variable)
#df <- df[which(complete.cases(df$p.Cresol.SO4.0) == TRUE),]

# Make categorical variables
input_data <- df1
out <- list()
for(i in seq_along(input_data)){
  mdn <- lapply(input_data, 2, FUN = median) # find the median for each column
  out[[i]] <- ifelse(input_data[,i] <= mdn[i], "below", "above") # assign below if value is less than or equal to median
}
res <- do.call("cbind",out)
colnames(res) <- paste0(colnames(input_data), "_mdn")
rownames(res) <- rownames(input_data)

# Turn to a data frame and rename. Then turn variables into factors as they should be.
str(res[,1:15])
df2 <- as.data.frame(res)
str(df2[,1:15])

# Factor
df3 <- df2 # back up
df3[] <- lapply(df3, function(x) {
  if(is.character(x)) as.factor(x) else x
})
# Check the levels. R will factor alphabetically but we want the "control" factor to be the below level
str(df3[,1:15])

# Merge categorical data with numeric data
df4 <- merge(df, df3.5, by = 0)
rownames(df4) <- df4[,1]
df4 <- df4[,-1]

# Assign 'Description' (subject ID in XXXX terms) to row names
rownames(md2) <- md2$Description
```

```{r Save ln biocrates data}
write.csv(md2, "../../data/processed/microbiome/phyloseq_inputs/lnBiocrates_mdn_sampledata_211108.csv")
saveRDS(md2, "../../data/processed/microbiome/phyloseq_inputs/lnBiocrates_mdn_sampledata_211108.rds")
```

### Make Phyloseq object - log biocrates BD1

```{r Make Phyloseq object- log biocrates BD1}
# Load files (metadata, count table, taxonomy table, tree)
# Metadata
metadata_a <- readRDS("../../data/processed/microbiome/phyloseq_inputs/lnBiocrates_mdn_sampledata_211108.rds")
# Count table
ASVr <- readRDS("../../data/processed/microbiome/phyloseq_inputs/merged_table_dada2_no_mito_chloro_highseqreps_0.2rm_subjectID.rds")
# Taxonomy table
taxtable <- readRDS("../../data/processed/microbiome/phyloseq_inputs/merged-taxonomy-dada2-retrained.rds")
# Phylogenetic Tree, rooted
treeR <- read_qza(file = "../../data/raw/from_MaryKable/tmao-tree-files/rooted-tree-dada2_no-mito-chloro_0.2rm_TMAOsamps_mc1.qza",  tmp = "Q2Rtemp", rm = F)

# Create phyloseq object
PSOlogBioc <- phyloseq(
  otu_table(ASVr, taxa_are_rows = T), 
  phy_tree(treeR$data), 
  tax_table(as.data.frame(taxtable) %>% dplyr::select(-Confidence) %>% column_to_rownames("Feature.ID") %>% as.matrix()), #moving the taxonomy to the way phyloseq wants it
  sample_data(metadata_a)
)
print(PSOlogBioc)

# Save the phyloseq object
save(PSOlogBioc, file = "../../data/processed/microbiome/phyloseq_objects/PSOlogBioc_211108.RData")
```

@@

```{r Biocrates BD1 log transformed phyloseq preprocess}
# Load Mary's original memtadata file... that we converted into a csv in the above chunk. 
md <- readRDS("../../data/processed/microbiome/phyloseq_inputs/FL100_mapping_file_complete_highest-seq-reps_nocontrols_min_withAlphaStatistics_211104.rds")
rownames(md) <- md[,"Description"]

>>>>

1) Processed QIIME2 inputs in R .rds and .csv formats
2) Phyloseq object with TMAO related phenotypes [PSOtmao]

**Updates:**
On 01-19-2022 I updated the chunks, "Original Data from Mary" and "Make Phyloseq object- specific TMAO variables". When MK was reading my manuscript, she noticed that I have some entries in the taxonomy table that say NA and some that say "g__" or "f__". While these may have different meanings, she suggests simply changing all cases to NA. (Another solution, you could manually differentaite "g__" to "g__ASV1" so that the glomming isn't wrong). I added code (from MK) to consistently change the NAs. Then, I created a new phyloseq object, PSO_tmao_NA. These will be added to the project branch named "update-taxonomy-table" in this GitHub project. The branches will eventually be merged to one.

@@

@@

```{r TMAO phyloseq preprocess}
# Load Mary's original memtadata file... that we converted into a csv in the above chunk. 
md <- readRDS("../../data/processed/microbiome/phyloseq_inputs/FL100_mapping_file_complete_highest-seq-reps_nocontrols_min_withAlphaStatistics_211104.rds")
rownames(md) <- md[,"Description"]
